Characterizing Signal Loss, Error, and Bias in the 21cm Reionization \\Power Spectrum: A Revised Study of PAPER-64

### 2018-06-22 PAPER.tex
small edits up until the end of the paper. Changed all i.e.'s to i.e.,'s

Armed with this information,
%Using what we've learned about the eigenspectrum, 
we can tweak the covariance in a simple way to suppress foregrounds and yield minimal signal loss. Recall that our toy model foreground 
% JEA - modes need not be sinusoids; the point is that there is one outer product describing this in the covariance
%is a sinusoid, so it 
can be perfectly described by a single eigenmode. Using the full dataset's (foreground plus EoR signal) empirical covariance, we can 
project out the zeroth eigenmode and 
%then flatten the rest of the spectrum to have eigenvalues of 1, thereby down-weighting the foreground-dominated mode more than the rest of the modes. Hence, we are changing the EoR-dominating part of the spectrum to be less coupled to the data, limiting the amount of over-fitting that can happen for those modes (i.e. only allowing over-fitting to occur for the first mode).
then take the remaining covariance to be the identity matrix.  
This decouples the covariance from the data for the EoR modes.  The resulting power spectrum estimate for this case is shown in the right plot of Figure \ref{fig:toy_sigloss3}. 
In this case we recover the EoR signal, demonstrating that if we can disentangle the foreground-dominated modes and EoR-dominated modes, we can suppress
% down-weight them 
foregrounds with negligible signal loss.

@@

% JEA - Death to eigenspectra!
%The resulting eigenspectrum as compared to the green curve (FG + EoR) in Figure \ref{fig:toy_sigloss2} is shown in Figure \ref{fig:toy_sigloss15}. 
With fringe-rate filtering resulting in fewer independent modes, it becomes more difficult for the empirical covariance to estimate the true covariance matrix of the fringe-rate filtered data. We can quantify this effect by evaluating a convergence metric $\varepsilon(\Chat)$ for the empirical covariance, which we define as
\begin{equation}
\label{eq:converge}
\varepsilon (\Chat) \equiv \sqrt{\frac{\sum_{ij} (\widehat{C}_{ij} - {C}_{ij})^2}{\sum_{ij} {C}_{ij}^2}},
\end{equation}
where $\C$ is the true covariance matrix. To compute this metric, we draw different numbers of realizations (different draws of Gaussian noise) of our toy model EoR measurement, $\textbf{x}_{\rm EoR}$, and take their ensemble average. We then compare this to the ``true" covariance, which in our simulation is set to be the empirical covariance after a large number ($500$) of realizations. As shown in Figure \ref{fig:toy_sigloss16}, we perform this computation for a range of total independent ensemble realizations (horizontal axis) and number of independent samples in the data following fringe-rate filtering (different colors). With more independent time samples (i.e. more realizations) in the data, one converges to the true fringe-rate filtered covariance more quickly.

@@

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{plots/toy_sigloss16.pdf}
	\caption{The convergence level, as defined by Equation \eqref{eq:converge}, of empirically estimated covariances of mock EoR signals with different numbers of independent samples. In red, the mock EoR signal is comprised entirely of independent samples. Subsequent colors show time-averaged signals. As the number of realizations increases, we see that the empirical covariances approach the true covariances. With more independent samples, the quicker an empirical covariance converges (i.e. the quicker it decouples from the data), and the less signal loss we would expect to result.}
	\label{fig:toy_sigloss16}
\end{figure}

@@

As a first test, we model the covariance matrix of EoR as a proof of concept that if perfect models are known, signal loss can be 
avoided. We know that our simulated EoR signal should have a covariance matrix that mimics the identity matrix, with its 
variance encoded along the diagonal. We model $\textbf{C}_{\rm EoR}$ as such (i.e. the identity), instead of computing it based on $\textbf{x}
_{\rm EoR}$ itself. Next, we add $\textbf{C}_{\rm EoR} + \widehat{\textbf{C}}_{\rm FG}$ (where $\widehat{\textbf{C}}_{\rm FG} = \langle\textbf{x}_{\rm FG}
\textbf{x}_{\rm FG}^{\dagger}\rangle_{t}$) to obtain a final $\widehat{\textbf{C}}$ to use in weighting. In Figure \ref{fig:toy_sigloss8} (upper 
left), we see that there is negligible signal loss. This is because by modeling $\textbf{C}_{\rm EoR}$, we avoid over-fitting EoR fluctuations in the data that our model doesn't know about (but an empirically derived $\widehat{\textbf{C}}_{\rm EoR}$ would). 
%This is also shown by comparing the (steeper) green and (flatter) red curves in Figure \ref{fig:toy_sigloss14}. 
In practice such a weighting option is not feasible, as it is difficult to model $\textbf{C}_{\rm EoR}$, and $\widehat{\textbf{C}}_{\rm FG}$ is unknown because we do not know how to separate out the foregrounds from the EoR in our data.

@@

The last regularization scheme we are highlighting here is setting $\widehat{\textbf{C}} \equiv \widehat{\textbf{C}} \circ \textbf{I}$ (element-wise multiplication), or inverse variance weighting (i.e. keeping only the diagonal elements of $\widehat{\textbf{C}}$). In the bottom right 
panel of Figure \ref{fig:toy_sigloss8}, we see that this method does not down-weight the foregrounds at all --- this regularization altered $\widehat{\textbf{C}}$ in a way where it is no longer coupled to \textit{any} of the empirically estimated eigenmodes, including the FG-dominated one. For this toy model, 
our foregrounds are spread out in frequency and therefore have non-negligible frequency-frequency correlations. Multiplying by 
the identity, element-wise, results in a diagonal matrix, meaning we do not have any correlation information. Because of this, we do a poor job 
suppressing the foreground. But because we decoupled the whole eigenspectrum from the data, we also avoid signal loss. Although this method did not successfully recover the EoR signal for this particular simulation, it is important that we show that there 
are many options for estimating a covariance matrix, and some may down-weight certain eigenmodes more effectively than others based on the spectral nature 
of the components in a dataset. 
%One may imagine a situation where a particular systematic is contained to an isolated 
%frequency (such as radio frequency interference or crosstalk). In such a case, preserving only the diagonal elements of $
%\widehat{\textbf{C}}$ would be an effective way of removing this contamination.

@@

Suppose we have $N$ different measurements targeting the same quantity (for example, $N$ power spectrum measurements). 
Bootstrapping means that we form $N_{\rm boot}$ (often a large number) bootstraps, where each bootstrap is a random selection 
of the $N$ measurements. Bootstraps each have dimensions of $N$, and the values populated into each bootstrap are drawn 
from the original set of measurements with replacement (i.e. every $n^{th}$ slot in $N$ is filled randomly for each bootstrap). Next we take 
the mean of each bootstrap to collapse it from an array of length $N$ to a single number (we are interested in the mean statistic 
here, but any function of interest can be applied to each bootstrap as long as it's the same function for each one). The error (on 
the mean) is then computed as the standard deviation across all bootstraps.

@@

%\begin{figure}
%	\centering
%	\includegraphics[trim={0cm 0cm 0cm 0cm},width=\columnwidth]{plots/toy_error2.png}
%	\caption{The fraction of independent elements in one bootstrap sample as a function of the total number of elements in a 
%dataset (i.e. if $4$ out of $5$ elements are independent, this would yield a fraction of $0.8$). A fraction of $1.0$ means that all 
%elements in the bootstrap sample are independent, and therefore sensitivity is maximized (alternately, this axis can be thought 
%of as power spectrum sensitivity, where sensitivity increases moving upwards along the y-axis). Two bootstrapping methods are shown here --- sampling all elements with replacement (gray) and sampling only the last element with replacement (black). The former fails to maximize sensitivity since some elements end up being repeated due to random sampling.} 
%The `outlier' points lying at a fraction of $1.0$ at the far left of the plot occur in the off-chance that the last spot in the sample happens to be the one missing element instead of a repeated element --- this is more likely to happen with a small total number of elements but also results in a more dramatically different fraction as compared to the case of a repeated element in the last spot, hence why these points appear like outliers.}
%	\label{fig:toy_error2}
%\end{figure}

@@

The jackknife is a resampling technique in which a statistic (i.e. power spectrum) is computed in subsets of the data (\citealt{quenouille1949}; \citealt{tukey1958}). These 
subsets are then compared to reveal systematics. In this section we define two main tests --- the null test and the traditional 
jackknife --- and explain how a power spectrum detection must pass each. We then highlight how these tests can be used to 
help distinguish between different sources of bias.

\begin{itemize}
\item \textbf{Null Test}: A null test is a type of jackknife test that removes the astronomical signal from data in order to 
investigate underlying systematics (see \citet{keating_et_al2016} for examples from intensity mapping that are closely related to our current application). For example, one can 
divide data into two subsets by separating odd and even Julian dates, or the first half of the observing season from the second. 
Subtracting the two removes signal that is common to both subsets, including foregrounds and the EoR signal. The resulting power 
spectrum should be consistent with thermal noise estimates; if it is not, it suggests the presence of a systematic that differs 
from one of the data subsets to the other (i.e. doesn't get subtracted perfectly). 
\item \textbf{Traditional Jackknife}: In a broader sense, it is important to perform many jackknife tests in order to instill 
confidence in a final result. A stable result must be steadfast throughout all jackknives no matter how the data is sliced. 
Jackknives can be taken along several different axes --- for example, one could start with a full dataset, and compute a new 
power spectrum every time as a day of data is removed, or a baseline is removed. This type of jackknife would reveal bias 
present only at certain LSTs (such as a foreground source), for example, or misbehaving baselines.
\end{itemize}

@@

\begin{figure}
	\centering
	\includegraphics[trim={0cm 3cm 0cm 3cm},width=\columnwidth]{plots/ant_layout_aspect.pdf}
	\caption{The PAPER-64 antenna layout. We use only $10$ of the $30$ m East/West baselines for the analysis in this 
paper (i.e. a subset of the shortest horizontal spacings).}
	\label{fig:ant_layout}
\end{figure}

@@

%In the limit where no significant cross-correlations between the data and the injected signals are found, this will recover the amount of injected power spectrum remaining after the covariance weighting operation (i.e. the amount of background EoR power, $P_{\rm eor}$, not removed by the foreground down-weighting procedure). In other words, we make the anzatz that

@@

{\bf Large injection:}
When the injected signal is much larger than the measured power spectrum, the data-only components can 
be neglected as many orders of magnitude smaller. We include a description of this regime for completeness in our discussion, but note that the upper limits that we compute are typically not determined by simulations in this regime (i.e. in using an empirical weighting scheme we've assumed the data to be dominated by foregrounds rather than the cosmological signal).  However, it is useful as a check of our system in a relatively simple case. As we can see from Figure \ref{fig:sigloss_terms}, the cross-terms (red) are small in comparison to the signal-only term (green). Here only does the signal-only term used in  \citetalias{ali_et_al2015} dominate the total power output. We again see that, in the empirical inverse covariance weighted case, the cross-terms behave as noise (positive and negative fluctuations around zero mean). This is for the same reason as at small injections --- here $\widehat{\textbf{C}}_{r}$ is dominated by the signal $\textbf{e}$. The cross-correlation can again be modeled as a symmetric distribution of zero mean and width $\sqrt{\widehat{\textbf{P}}_e}\sqrt{\widehat{\textbf{P}}_x}$.

{\bf In between:}
When the injected signal is of a similar amplitude to the data by itself, the situation becomes less straightforward. We see that 
the weighted injected power spectrum component mirrors the input power indicating little loss (i.e. the green curve follows the dotted black line), eventually 
departing from unity when the injected amplitude is well above the level of the data power spectrum. However, 
in this regime the cross-term (red) has nearly the same amplitude, but with a negative sign. As explained below, this negativity is the result of cross-correlating inverse covariance weighted terms.  This negative component drives down the $\widehat{P}_{\rm out}$ estimator (black). We note that in \citetalias{ali_et_al2015}, signal loss was computed by only looking at the second term in Equation \eqref{eq:crossterm} (green), which incorrectly implies no loss at the data-only power spectrum level. Ignoring the effect of the negative power from the cross-terms is the main reason for under-estimating power spectrum limits in \citetalias{ali_et_al2015}.

@@

\noindent where we can neglect the first term because $\textbf{x}$ is small (i.e. the large negative cross-term power in the left panel of Figure \ref{fig:sigloss_terms} occurs when the injected amplitude surpasses the level of the data-only power spectrum).  Without loss of generality, we will assume
an eigenbasis of $\textbf{e}$, so that $\langle 
\textbf{ee}^{\dagger} \rangle_{t}$ is diagonal. The middle 
two terms, however, can have power in their off-diagonal terms due to the fact that, when averaging over a finite
ensemble, $\langle\textbf{xe}^\dagger\rangle_t$ is not zero.  As shown in Appendix C of \citet{parsons_et_al2014}%\footnote{This same appendix also gives a now-prescient warning about signal loss and the dangers of noise and sample variance in inverse covariance.}
, to leading order the inversion of a diagonal-dominant matrix like $\widehat{\textbf{C}}_{r}$ (from $\langle 
\textbf{ee}^{\dagger} \rangle_{t}$) with smaller
off-diagonal terms results in a new diagonal-dominant matrix with negative off-diagonal terms. These off-diagonal
terms depend on both $\textbf{x}$ and $\textbf{e}$. Then, when $\widehat{\textbf{C}}^{-1}_{r}$ is multiplied into $\textbf{x}$,
the result is a vector that is similar to $\textbf{x}$ but
contains a residual correlation to $\textbf{e}$ from the off-diagonal components of $\widehat{\textbf{C}}^{-1}_{r}$. The
correlation is negative because the product $\widehat{\textbf{C}}_r^{-1}\textbf{x}$ effectively squares the $\textbf{x}$-dependence
of the off-diagonal terms in $\widehat{\textbf{C}}^{-1}_{r}$ while retaining the negative sign that arose from the inversion
of a diagonal-dominant matrix.
% XXX [ARP: we may want to write this all out explicitly in an appendix]

@@

One issue to address is how one incorporates the randomness of $\widehat{P}_{\rm out}$ into our signal loss corrections. A different realization of the mock EoR signal is injected with each bootstrap run, causing the output to vary in three ways ---  there is noise variation from the bootstraps, there is cosmic variation from generating multiple realizations of the mock EoR signal, and there is a variation caused by whether the injected signal looks more or less ``like'' the data (i.e. how much coupling there is, which affects how much loss results).

@@

We saw in Section \ref{sec:otherweight} how limiting the number of down-weighted eigenmodes (i.e. flattening out part of the 
eigenspectrum and effectively decoupling the lowest-valued eigenmodes, which are typically EoR-dominated, from the data) can help minimize signal loss. We experiment with this idea on PAPER-64 data, dialing the number of modes 
that are down-weighted from zero (which is equivalent to identity-weighting, or the uniform-weighted case) to $21$ (which is the full inverse 
covariance estimator). The power spectrum results for one $k$-value, both before and after signal loss 
estimation, are shown in the top panel in Figure \ref{fig:sigloss_modeloop}. We see that the amount of signal loss increases as weighting 
becomes more aggressive (dashed red). In other words, more EoR-dominated fluctuations are being overfit and 
subtracted as more modes are down-weighted. We also find that the power spectrum upper limit, post-signal loss estimation, 
increases with the number of down-weighted modes (solid red). The more modes we use in down-weighting, the stronger the coupling between the weighting and the data, and the greater the error we have in estimating the power spectrum. \citet{switzer_et_al2013} took a similar approach in determining the optimal number of modes to down-weight in GBT data, finding similar trends and noting that removing too few modes is limited by residual foregrounds and removing too many modes is limited by large error bars and signal loss.

@@

\begin{figure*}
	\centering
	\includegraphics[width=1\textwidth]{plots/sigloss_modeloop_2panel.pdf}
	\caption{Power spectra $2\sigma$ upper limits for $k=0.393$\,$h$ Mpc$^{-1}$ for fringe-rate filtered PAPER-64 data. Top: Values 
are shown before (dashed red) and after (solid red) signal loss estimation via our signal injection framework as a function of number of eigenmodes of $\widehat{\textbf{C}}$ that 
are down-weighted. This regularization knob is tuned from $0$ modes on the left (i.e. unweighted) to $21$ modes on the right (i.e. the full inverse 
covariance estimator). $\sim4$ orders of magnitude of signal loss results when using empirically estimated inverse covariance weighting. Bottom: Power spectrum upper limits before (dashed red) and after (solid red) signal loss estimation as a function of identity added to the empirical covariance. This regularization knob is tuned from $\gamma = 10^{-4}$ on the right (i.e. very little regularization) to $\gamma = 1$ on the left (see main text for the definition of $\gamma$). Also 
plotted in both panels for comparison are $2\sigma$ power spectrum upper limits for the uniform-weighted case (blue) and inverse variance 
weighted case (black); both are after signal loss estimation. Finally, a theoretical prediction for noise ($2\sigma$ error) is plotted 
as green. In the PAPER-64 analysis in this paper, we choose to use a regularization scheme of $\widehat{\textbf{C}}_{\rm eff} \equiv 0.09 \, $Tr($\widehat{\textbf{C}})\textbf{I} + \widehat{\textbf{C}}$ ($\gamma = 0.09$) as a simple example of regularization that minimizes loss, and note that the power spectrum limits using this type of regularization are roughly constant across a large range of values of $\gamma$.}
	\label{fig:sigloss_modeloop}
\end{figure*}

>>>>

Armed with this information,
%Using what we've learned about the eigenspectrum, 
we can tweak the covariance in a simple way to suppress foregrounds and yield minimal signal loss. Recall that our toy model foreground 
% JEA - modes need not be sinusoids; the point is that there is one outer product describing this in the covariance
%is a sinusoid, so it 
can be perfectly described by a single eigenmode. Using the full dataset's (foreground plus EoR signal) empirical covariance, we can 
project out the zeroth eigenmode and 
%then flatten the rest of the spectrum to have eigenvalues of 1, thereby down-weighting the foreground-dominated mode more than the rest of the modes. Hence, we are changing the EoR-dominating part of the spectrum to be less coupled to the data, limiting the amount of over-fitting that can happen for those modes (i.e., only allowing over-fitting to occur for the first mode).
then take the remaining covariance to be the identity matrix.  
This decouples the covariance from the data for the EoR modes.  The resulting power spectrum estimate for this case is shown in the right plot of Figure \ref{fig:toy_sigloss3}. 
In this case we recover the EoR signal, demonstrating that if we can disentangle the foreground-dominated modes and EoR-dominated modes, we can suppress
% down-weight them 
foregrounds with negligible signal loss.

@@

% JEA - Death to eigenspectra!
%The resulting eigenspectrum as compared to the green curve (FG + EoR) in Figure \ref{fig:toy_sigloss2} is shown in Figure \ref{fig:toy_sigloss15}. 
With fringe-rate filtering resulting in fewer independent modes, it becomes more difficult for the empirical covariance to estimate the true covariance matrix of the fringe-rate filtered data. We can quantify this effect by evaluating a convergence metric $\varepsilon(\Chat)$ for the empirical covariance, which we define as
\begin{equation}
\label{eq:converge}
\varepsilon (\Chat) \equiv \sqrt{\frac{\sum_{ij} (\widehat{C}_{ij} - {C}_{ij})^2}{\sum_{ij} {C}_{ij}^2}},
\end{equation}
where $\C$ is the true covariance matrix. To compute this metric, we draw different numbers of realizations (different draws of Gaussian noise) of our toy model EoR measurement, $\textbf{x}_{\rm EoR}$, and take their ensemble average. We then compare this to the ``true" covariance, which in our simulation is set to be the empirical covariance after a large number ($500$) of realizations. As shown in Figure \ref{fig:toy_sigloss16}, we perform this computation for a range of total independent ensemble realizations (horizontal axis) and number of independent samples in the data following fringe-rate filtering (different colors). With more independent time samples (i.e., more realizations) in the data, one converges to the true fringe-rate filtered covariance more quickly.

@@

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{plots/toy_sigloss16.pdf}
	\caption{The convergence level, as defined by Equation \eqref{eq:converge}, of empirically estimated covariances of mock EoR signals with different numbers of independent samples. In red, the mock EoR signal is comprised entirely of independent samples. Subsequent colors show time-averaged signals. As the number of realizations increases, we see that the empirical covariances approach the true covariances. With more independent samples, the quicker an empirical covariance converges (i.e., the quicker it decouples from the data), and the less signal loss we would expect to result.}
	\label{fig:toy_sigloss16}
\end{figure}

@@

As a first test, we model the covariance matrix of EoR as a proof of concept that if perfect models are known, signal loss can be 
avoided. We know that our simulated EoR signal should have a covariance matrix that mimics the identity matrix, with its 
variance encoded along the diagonal. We model $\textbf{C}_{\rm EoR}$ as such (i.e., the identity), instead of computing it based on $\textbf{x}
_{\rm EoR}$ itself. Next, we add $\textbf{C}_{\rm EoR} + \widehat{\textbf{C}}_{\rm FG}$ (where $\widehat{\textbf{C}}_{\rm FG} = \langle\textbf{x}_{\rm FG}
\textbf{x}_{\rm FG}^{\dagger}\rangle_{t}$) to obtain a final $\widehat{\textbf{C}}$ to use in weighting. In Figure \ref{fig:toy_sigloss8} (upper 
left), we see that there is negligible signal loss. This is because by modeling $\textbf{C}_{\rm EoR}$, we avoid over-fitting EoR fluctuations in the data that our model doesn't know about (but an empirically derived $\widehat{\textbf{C}}_{\rm EoR}$ would). 
%This is also shown by comparing the (steeper) green and (flatter) red curves in Figure \ref{fig:toy_sigloss14}. 
In practice such a weighting option is not feasible, as it is difficult to model $\textbf{C}_{\rm EoR}$, and $\widehat{\textbf{C}}_{\rm FG}$ is unknown because we do not know how to separate out the foregrounds from the EoR in our data.

@@

The last regularization scheme we are highlighting here is setting $\widehat{\textbf{C}} \equiv \widehat{\textbf{C}} \circ \textbf{I}$ (element-wise multiplication), or inverse variance weighting (i.e., keeping only the diagonal elements of $\widehat{\textbf{C}}$). In the bottom right 
panel of Figure \ref{fig:toy_sigloss8}, we see that this method does not down-weight the foregrounds at all --- this regularization altered $\widehat{\textbf{C}}$ in a way where it is no longer coupled to \textit{any} of the empirically estimated eigenmodes, including the FG-dominated one. For this toy model, 
our foregrounds are spread out in frequency and therefore have non-negligible frequency-frequency correlations. Multiplying by 
the identity, element-wise, results in a diagonal matrix, meaning we do not have any correlation information. Because of this, we do a poor job 
suppressing the foreground. But because we decoupled the whole eigenspectrum from the data, we also avoid signal loss. Although this method did not successfully recover the EoR signal for this particular simulation, it is important that we show that there 
are many options for estimating a covariance matrix, and some may down-weight certain eigenmodes more effectively than others based on the spectral nature 
of the components in a dataset. 
%One may imagine a situation where a particular systematic is contained to an isolated 
%frequency (such as radio frequency interference or crosstalk). In such a case, preserving only the diagonal elements of $
%\widehat{\textbf{C}}$ would be an effective way of removing this contamination.

@@

Suppose we have $N$ different measurements targeting the same quantity (for example, $N$ power spectrum measurements). 
Bootstrapping means that we form $N_{\rm boot}$ (often a large number) bootstraps, where each bootstrap is a random selection 
of the $N$ measurements. Bootstraps each have dimensions of $N$, and the values populated into each bootstrap are drawn 
from the original set of measurements with replacement (i.e., every $n^{th}$ slot in $N$ is filled randomly for each bootstrap). Next we take 
the mean of each bootstrap to collapse it from an array of length $N$ to a single number (we are interested in the mean statistic 
here, but any function of interest can be applied to each bootstrap as long as it's the same function for each one). The error (on 
the mean) is then computed as the standard deviation across all bootstraps.

@@

%\begin{figure}
%	\centering
%	\includegraphics[trim={0cm 0cm 0cm 0cm},width=\columnwidth]{plots/toy_error2.png}
%	\caption{The fraction of independent elements in one bootstrap sample as a function of the total number of elements in a 
%dataset (i.e., if $4$ out of $5$ elements are independent, this would yield a fraction of $0.8$). A fraction of $1.0$ means that all 
%elements in the bootstrap sample are independent, and therefore sensitivity is maximized (alternately, this axis can be thought 
%of as power spectrum sensitivity, where sensitivity increases moving upwards along the y-axis). Two bootstrapping methods are shown here --- sampling all elements with replacement (gray) and sampling only the last element with replacement (black). The former fails to maximize sensitivity since some elements end up being repeated due to random sampling.} 
%The `outlier' points lying at a fraction of $1.0$ at the far left of the plot occur in the off-chance that the last spot in the sample happens to be the one missing element instead of a repeated element --- this is more likely to happen with a small total number of elements but also results in a more dramatically different fraction as compared to the case of a repeated element in the last spot, hence why these points appear like outliers.}
%	\label{fig:toy_error2}
%\end{figure}

@@

The jackknife is a resampling technique in which a statistic (i.e., power spectrum) is computed in subsets of the data (\citealt{quenouille1949}; \citealt{tukey1958}). These 
subsets are then compared to reveal systematics. In this section we define two main tests --- the null test and the traditional 
jackknife --- and explain how a power spectrum detection must pass each. We then highlight how these tests can be used to 
help distinguish between different sources of bias.

\begin{itemize}
\item \textbf{Null Test}: A null test is a type of jackknife test that removes the astronomical signal from data in order to 
investigate underlying systematics (see \citet{keating_et_al2016} for examples from intensity mapping that are closely related to our current application). For example, one can 
divide data into two subsets by separating odd and even Julian dates, or the first half of the observing season from the second. 
Subtracting the two removes signal that is common to both subsets, including foregrounds and the EoR signal. The resulting power 
spectrum should be consistent with thermal noise estimates; if it is not, it suggests the presence of a systematic that differs 
from one of the data subsets to the other (i.e., doesn't get subtracted perfectly). 
\item \textbf{Traditional Jackknife}: In a broader sense, it is important to perform many jackknife tests in order to instill 
confidence in a final result. A stable result must be steadfast throughout all jackknives no matter how the data is sliced. 
Jackknives can be taken along several different axes --- for example, one could start with a full dataset, and compute a new 
power spectrum every time as a day of data is removed, or a baseline is removed. This type of jackknife would reveal bias 
present only at certain LSTs (such as a foreground source), for example, or misbehaving baselines.
\end{itemize}

@@

\begin{figure}
	\centering
	\includegraphics[trim={0cm 3cm 0cm 3cm},width=\columnwidth]{plots/ant_layout_aspect.pdf}
	\caption{The PAPER-64 antenna layout. We use only $10$ of the $30$ m East/West baselines for the analysis in this 
paper (i.e., a subset of the shortest horizontal spacings).}
	\label{fig:ant_layout}
\end{figure}

@@

%In the limit where no significant cross-correlations between the data and the injected signals are found, this will recover the amount of injected power spectrum remaining after the covariance weighting operation (i.e., the amount of background EoR power, $P_{\rm eor}$, not removed by the foreground down-weighting procedure). In other words, we make the anzatz that

@@

{\bf Large injection:}
When the injected signal is much larger than the measured power spectrum, the data-only components can 
be neglected as many orders of magnitude smaller. We include a description of this regime for completeness in our discussion, but note that the upper limits that we compute are typically not determined by simulations in this regime (i.e., in using an empirical weighting scheme we've assumed the data to be dominated by foregrounds rather than the cosmological signal).  However, it is useful as a check of our system in a relatively simple case. As we can see from Figure \ref{fig:sigloss_terms}, the cross-terms (red) are small in comparison to the signal-only term (green). Here only does the signal-only term used in  \citetalias{ali_et_al2015} dominate the total power output. We again see that, in the empirical inverse covariance weighted case, the cross-terms behave as noise (positive and negative fluctuations around zero mean). This is for the same reason as at small injections --- here $\widehat{\textbf{C}}_{r}$ is dominated by the signal $\textbf{e}$. The cross-correlation can again be modeled as a symmetric distribution of zero mean and width $\sqrt{\widehat{\textbf{P}}_e}\sqrt{\widehat{\textbf{P}}_x}$.

{\bf In between:}
When the injected signal is of a similar amplitude to the data by itself, the situation becomes less straightforward. We see that 
the weighted injected power spectrum component mirrors the input power indicating little loss (i.e., the green curve follows the dotted black line), eventually 
departing from unity when the injected amplitude is well above the level of the data power spectrum. However, 
in this regime the cross-term (red) has nearly the same amplitude, but with a negative sign. As explained below, this negativity is the result of cross-correlating inverse covariance weighted terms.  This negative component drives down the $\widehat{P}_{\rm out}$ estimator (black). We note that in \citetalias{ali_et_al2015}, signal loss was computed by only looking at the second term in Equation \eqref{eq:crossterm} (green), which incorrectly implies no loss at the data-only power spectrum level. Ignoring the effect of the negative power from the cross-terms is the main reason for under-estimating power spectrum limits in \citetalias{ali_et_al2015}.

@@

\noindent where we can neglect the first term because $\textbf{x}$ is small (i.e., the large negative cross-term power in the left panel of Figure \ref{fig:sigloss_terms} occurs when the injected amplitude surpasses the level of the data-only power spectrum).  Without loss of generality, we will assume
an eigenbasis of $\textbf{e}$, so that $\langle 
\textbf{ee}^{\dagger} \rangle_{t}$ is diagonal. The middle 
two terms, however, can have power in their off-diagonal terms due to the fact that, when averaging over a finite
ensemble, $\langle\textbf{xe}^\dagger\rangle_t$ is not zero.  As shown in Appendix C of \citet{parsons_et_al2014}%\footnote{This same appendix also gives a now-prescient warning about signal loss and the dangers of noise and sample variance in inverse covariance.}
, to leading order the inversion of a diagonal-dominant matrix like $\widehat{\textbf{C}}_{r}$ (from $\langle 
\textbf{ee}^{\dagger} \rangle_{t}$) with smaller
off-diagonal terms results in a new diagonal-dominant matrix with negative off-diagonal terms. These off-diagonal
terms depend on both $\textbf{x}$ and $\textbf{e}$. Then, when $\widehat{\textbf{C}}^{-1}_{r}$ is multiplied into $\textbf{x}$,
the result is a vector that is similar to $\textbf{x}$ but
contains a residual correlation to $\textbf{e}$ from the off-diagonal components of $\widehat{\textbf{C}}^{-1}_{r}$. The
correlation is negative because the product $\widehat{\textbf{C}}_r^{-1}\textbf{x}$ effectively squares the $\textbf{x}$-dependence
of the off-diagonal terms in $\widehat{\textbf{C}}^{-1}_{r}$ while retaining the negative sign that arose from the inversion
of a diagonal-dominant matrix.
% XXX [ARP: we may want to write this all out explicitly in an appendix]

@@

One issue to address is how one incorporates the randomness of $\widehat{P}_{\rm out}$ into our signal loss corrections. A different realization of the mock EoR signal is injected with each bootstrap run, causing the output to vary in three ways ---  there is noise variation from the bootstraps, there is cosmic variation from generating multiple realizations of the mock EoR signal, and there is a variation caused by whether the injected signal looks more or less ``like'' the data (i.e., how much coupling there is, which affects how much loss results).

@@

We saw in Section \ref{sec:otherweight} how limiting the number of down-weighted eigenmodes (i.e., flattening out part of the 
eigenspectrum and effectively decoupling the lowest-valued eigenmodes, which are typically EoR-dominated, from the data) can help minimize signal loss. We experiment with this idea on PAPER-64 data, dialing the number of modes 
that are down-weighted from zero (which is equivalent to identity-weighting, or the uniform-weighted case) to $21$ (which is the full inverse 
covariance estimator). The power spectrum results for one $k$-value, both before and after signal loss 
estimation, are shown in the top panel in Figure \ref{fig:sigloss_modeloop}. We see that the amount of signal loss increases as weighting 
becomes more aggressive (dashed red). In other words, more EoR-dominated fluctuations are being overfit and 
subtracted as more modes are down-weighted. We also find that the power spectrum upper limit, post-signal loss estimation, 
increases with the number of down-weighted modes (solid red). The more modes we use in down-weighting, the stronger the coupling between the weighting and the data, and the greater the error we have in estimating the power spectrum. \citet{switzer_et_al2013} took a similar approach in determining the optimal number of modes to down-weight in GBT data, finding similar trends and noting that removing too few modes is limited by residual foregrounds and removing too many modes is limited by large error bars and signal loss.

@@

\begin{figure*}
	\centering
	\includegraphics[width=1\textwidth]{plots/sigloss_modeloop_2panel.pdf}
	\caption{Power spectra $2\sigma$ upper limits for $k=0.393$\,$h$ Mpc$^{-1}$ for fringe-rate filtered PAPER-64 data. Top: Values 
are shown before (dashed red) and after (solid red) signal loss estimation via our signal injection framework as a function of number of eigenmodes of $\widehat{\textbf{C}}$ that 
are down-weighted. This regularization knob is tuned from $0$ modes on the left (i.e., unweighted) to $21$ modes on the right (i.e., the full inverse 
covariance estimator). $\sim4$ orders of magnitude of signal loss results when using empirically estimated inverse covariance weighting. Bottom: Power spectrum upper limits before (dashed red) and after (solid red) signal loss estimation as a function of identity added to the empirical covariance. This regularization knob is tuned from $\gamma = 10^{-4}$ on the right (i.e., very little regularization) to $\gamma = 1$ on the left (see main text for the definition of $\gamma$). Also 
plotted in both panels for comparison are $2\sigma$ power spectrum upper limits for the uniform-weighted case (blue) and inverse variance 
weighted case (black); both are after signal loss estimation. Finally, a theoretical prediction for noise ($2\sigma$ error) is plotted 
as green. In the PAPER-64 analysis in this paper, we choose to use a regularization scheme of $\widehat{\textbf{C}}_{\rm eff} \equiv 0.09 \, $Tr($\widehat{\textbf{C}})\textbf{I} + \widehat{\textbf{C}}$ ($\gamma = 0.09$) as a simple example of regularization that minimizes loss, and note that the power spectrum limits using this type of regularization are roughly constant across a large range of values of $\gamma$.}
	\label{fig:sigloss_modeloop}
\end{figure*}

In addition to our thermal noise prediction (green) and uniform-weighted power spectrum limit (blue), one additional horizontal line is shown in Figure \ref{fig:sigloss_modeloop} 
in both panels and represents a third regularization technique. This line (black) denotes the power spectrum value, post-signal loss estimation, for inverse variance weighting (multiplying an identity 
matrix element-wise to $\widehat{\textbf{C}}$). This result is single-valued and not a function of the horizontal axis. We see that all three regularization schemes shown (solid red top panel, solid red bottom panel, black) perform similarly at 
their best (i.e., when $\sim2$-$3$ eigenmodes are down-weighted in the case of the top panel's solid red curve). However, for the remainder of this paper, we choose to use the weighting option of $\widehat{\textbf{C}} \equiv \widehat{\textbf{C}} + 0.09 \,$Tr($\widehat{\textbf{C}})\textbf{I}$, or $\gamma = 0.09$, which we will denote as $\widehat{\textbf{C}}_{\rm eff}$. We choose this weighting scheme merely as a simple example of regularizing PAPER-64 covariances, noting that the power spectrum upper limit remains roughly constant for a broad range of values of $\gamma$. We also note that all of the regularizations described here perform similarly to the uniformly weighted case. Thus, in Kolopanis et al. (\textit{in prep.}), uniform weightings are used to produce straightforward power spectrum limits that do not suffer from loss.
